---
title: "Hockey Shot Analysis"
format: gfm
---
```{r}
#| output: false
#| echo: false

library(tidyverse)
library(sportyR)
library(here)

shots_df <- read_csv(here("data/shots_2025.csv"), show_col_types = FALSE)

period_labels <- c(
  "1" = "Period 1\nVAN 3 - COL 2",
  "2" = "Period 2\nVAN 6 - COL 3",
  "3" = "Period 3\nFinal: VAN 8 - COL 6"
)

avs_vs_canucks <- shots_df |>
  filter(
    homeTeamCode == "COL",
    awayTeamCode == "VAN",
    game_id == 21188,
    period %in% c(1, 2, 3),
    event %in% c("SHOT", "GOAL")
  ) |>
  mutate(
    team_label = case_when(
      teamCode == "COL" ~ "Colorado",
      teamCode == "VAN" ~ "Vancouver",
      TRUE ~ teamCode
    ),
    event_label = case_when(
      event == "GOAL" ~ "Goal",
      event == "SHOT" ~ "Shot",
      TRUE ~ event
    ),
    period_lab = factor(
      as.character(period),
      levels = c("1", "2", "3"),
      labels = period_labels
    )
  ) |>
  filter(
    !is.na(arenaAdjustedXCord),
    !is.na(arenaAdjustedYCord)
  )

goal_df <- avs_vs_canucks |>
  filter(event == "GOAL")
```

## Data Introduction

- This project uses NHL play-by-play data from the 2025-2026 regular season.

- Each event includes information such as game ID, period, shot type, team context, and rink coordinates.

- Because the data include shot location, we can study where offenses generated chances on the ice.

- I included only shot-on-goal and goal events, and removed rows with missing shot type values.

## Questions of Interest

- How did shot location differ between the home and away teams in this game?

- Did shot patterns change from period to period as the game progressed?

```{r}
#| echo: false
team_colors <- c(
  "Colorado" = "#6F263D",
  "Vancouver" = "#00843D"
)


geom_hockey(league = "NHL") +
  geom_point(
    data = avs_vs_canucks,
    aes(
      x = arenaAdjustedXCord,
      y = arenaAdjustedYCord,
      color = team_label
    ),
    alpha = 0.75,
    size = 2.8
  ) +
  facet_wrap(~ period_lab) +
  scale_color_manual(values = team_colors) +
  labs(
    title = "Shots taken by each team",
    caption = "Vancouver at Colorado, Apr. 1, 2026",
    color = "Team"
  ) +
  theme_void(base_size = 18) +
  theme(
    plot.title = element_text(face = "bold", size = 22),
    plot.subtitle = element_text(size = 14),
    strip.text = element_text(face = "bold", size = 15),
    legend.position = "bottom"
  )
```

```{r}
#| echo: false
geom_hockey(league = "NHL") +
  geom_point(
    data = avs_vs_canucks,
    aes(x = arenaAdjustedXCord, y = arenaAdjustedYCord),
    color = "gray75",
    alpha = 0.45,
    size = 2
  ) +
  geom_point(
    data = goal_df,
    aes(
      x = arenaAdjustedXCord,
      y = arenaAdjustedYCord,
      color = team_label
    ),
    size = 4.5
  ) +
  facet_wrap(~ period_lab) +
  scale_color_manual(values = team_colors) +
  labs(
    title = "Goals highlighted over all shots",
    caption = "Vancouver at Colorado, Apr. 1, 2026",
    color = "Team"
  ) +
  theme_void(base_size = 18) +
  theme(
    plot.title = element_text(face = "bold", size = 22),
    plot.subtitle = element_text(size = 14),
    strip.text = element_text(face = "bold", size = 15),
    legend.position = "bottom"
  )
```


```{r}
#| echo: false
#| message: false
avs_vs_canucks |>
  count(period_lab, team_label, event_label) |>
  ggplot(aes(x = period_lab, y = n, fill = event_label)) +
  geom_col(position = "dodge") +
  facet_wrap(~ team_label) +
  scale_fill_manual(values = c("Shot" = "#9ecae1", "Goal" = "#f28e2b")) +
  theme_minimal(base_size = 18) +
  labs(
    title = "Shot events by period",
    caption = "Vancouver at Colorado, Apr. 1, 2026",
    color = "Team",
    x = NULL,
    y = "Events",
    fill = "Event Type"
  ) +
  theme(
    plot.title = element_text(face = "bold", size = 22),
    plot.subtitle = element_text(size = 14),
    strip.text = element_text(face = "bold", size = 16),
    axis.text.x = element_text(size = 11),
    legend.position = "bottom"
  )
```
